\tealeaves: an R package for modeling leaf temperature using energy budgets

### 2019-01-24 ms/ms.tex
initial commit

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage[labelfont=bf]{caption}
\usepackage{fancyhdr}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt} % remove line below header
  \lhead{}
  \chead{\textit{Leaf energy budgets in R}} % running head
  \rhead{}
\usepackage{footnote}
  \makesavenoteenv{tabular}
\usepackage{geometry}
  \geometry{letterpaper}
\usepackage{lineno}
\usepackage{natbib}
\usepackage[parfill]{parskip}
\usepackage{setspace}
  \newcommand{\stretchy}{1.5}
\usepackage{Sweave}
  \DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=2em,frame=single}
  \DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=2em,frame=single}

@@

\begin{document}
\SweaveOpts{concordance=TRUE}

@@

$^1$ Department of Botany, University of Hawai'i, Honolulu, Hawai'i 96822, USA \\
\\
Phone: +1 808-688-3478 \\
Email: cdmuir@hawaii.edu \\
\\
Funding: University of Hawai'i \\
\\
\section*{Abstract}

Plants must regulate leaf temperature to optimize photosynthesis, control water loss, and prevent damage caused by overheating or freezing. Physical models of leaf energy budgets calculate the energy fluxes and leaf temperatures for given a set leaf and environmental parameters. These models can provide deep insight into the variation in leaf form and function, but there are few computational tools available to use these models. Here I introduce a new R package called \tealeaves to make complex leaf energy budget models accessible to a broader array of plant scientists. This package enables novice users to start modeling leaf energy budgets quickly while allowing experts customize their parameter settings. The code is open source, freely available, and readily integrates with other R tools for scientific computing. This paper describes the current functionality of \tealeaves, but new features will be added in future releases. This software tool will advance new research on leaf thermal physiology to advance our understanding of basic and applied plant science.

@@

@@

<<Import-objects, echo = FALSE, results = hide>>=

  setwd("..")
  source("r/header.R")

  # import2ms()
  setwd("ms")

@

@@

@@

\begin{equation}
  \label{eq:energy_budget}
  R_\mathrm{abs} = S_\mathrm{r} + H + L
\end{equation}

@@

@@

R is a fully open source programming language for statistical computing that allows users to develop their own packages with new functions. \tealeaves~takes three sets of parameter inputs: leaf parameters, environmental parameters, and physical constants (see Table \ref{table:table_input}). The package provides reasonable defaults, but users can input new values to address their question, as I demonstrate in the next section. With one or more parameter sets, \tealeaves~uses the \code{uniroot} function in R base package \pkg{stats} to find the \tleaf~that balances the leaf energy budget (Equation \ref{eq:energy_budget}). It outputs the equilibrium \tleaf~and energy fluxes in a table for analysis and visualization.

@@

<<Example1, echo = TRUE, eval = FALSE, results = hide>>=

  library(tealeaves)

  # Default parameter inputs  
  leaf_par   <- make_leafpar()
  enviro_par <- make_enviropar()
  constants  <- make_constants()

  # Solve for T_leaf
  T_leaf <- tleaf(leaf_par, enviro_par, constants, 
                  quiet = TRUE, unitless = FALSE)

@

@@

<<Example2, echo = TRUE, eval = FALSE, results = hide>>=

  library(tealeaves)

  # Custom parameter inputs
  leaf_par   <- make_leafpar(
    replace = list(
      leafsize = set_units(c(0.0025, 0.025, 0.25), "m")
      )
  )
  enviro_par <- make_enviropar(
    replace = list(
      T_air = set_units(seq(275, 310, 5), "K")
      )
  )
  constants  <- make_constants()

  # Solve for T_leaf over a range of T_air
  T_leaves <- tleaves(leaf_par, enviro_par, constants, 
                      quiet = TRUE, unitless = FALSE)

@

>>>>

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage[labelfont=bf]{caption}
\usepackage{fancyhdr}
  \pagestyle{fancy}
  %\fancyhf{}
  \renewcommand{\headrulewidth}{0pt} % remove line below header
  \lhead{}
  \chead{\textit{Leaf energy budgets in R}} % running head
  \rhead{}
\usepackage{footnote}
  \makesavenoteenv{tabular}
\usepackage{geometry}
  \geometry{letterpaper}
\usepackage{lineno}
\usepackage{natbib}
\usepackage[parfill]{parskip}
\usepackage{setspace}
  \newcommand{\stretchy}{1.5}
\usepackage{Sweave}
  \DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=2em,frame=single}
  \DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=2em,frame=single}

@@

\begin{document}
\input{ms-concordance}

@@

$^1$ Department of Botany, University of Hawai'i, Honolulu, Hawai'i 96822, USA \\
Phone: +1 808-688-3478 \\
Email: cdmuir@hawaii.edu \\
\\
Funding: University of Hawai'i \\
\\
\section*{Abstract}

Make it easier for those working on plant ecophysiology to carry out advances studies of leaf temperature and its consequences.

@@

%reminder:
% free convection - radiator, diff in air density
% forced convection - fan, external pressure gradient

% reviewers:
% Tom Buckley
% Martjin Slot
% Adrienne Nicotra
% Andrea Leigh
% Bill Smith
% Okajima or Ichiro Terashima
% Joseph Stinazano
% Remko Duurmsma
% Schymanski SJ
% Sean Michelatz
% Paul Drake, Hugo de Boer, Erik Venkelaas

@@

%%%%% NOTES %%%%%

% According to Slot and Winter 2017 PCE: Above TOpt, both stomatal and biochemical factors become increasingly limiting to net photosynthesis (Sage & Kubien, 2007), but biochemical parameters ...

% Rogers et al NP - recommended better understanding of how leaf temperature affects photosynthesis in global change models. This could be a helpful tool for incorporating leaf temperature.

% Slot and Winter 2017 NP - found Topt near air temperature for all species. no specialist-generalist tradeoff.
% Sage and Kubien 2007

% history
% introduction of "boundary layer"
% Hofman 1955, 1956
% Hofmann, G., Die Thennodynamik der Taubildung (Ber. d. Deut. Wetterdienstes, 3, Nr. 18, Bad Kissingen, Germany, 45 pp., 1955)
% Hofmann, G., Planta, 47, 303-22 (1956)
% Raschke 1956. Raschke, K., Planta, 48,200-38 (1956)

% introduction of energy balance (as cited in Raschke 1960)
% Brown and Escombe. 1905. Proc. Roy. Soc. (London) B, 76, 29-111 (1905)
% Buttner 1934: Die Obertragung durch Leitung und Konvektion, Verdunstung und Strahlung in Bioklimatologie und Meteor% gie (Abh. preuss. Meteorol. Institut., 10, Nr. 5, 37 pp., 1934)
% Huber. 1935. Der Wiirmehaushalt der Pflanzen

% reviews
% Raschke 1960
% Gutschick 2016

% leaf energy 
% example papers:
% leaf size (theory?): 
%   Gates 1965
%   Drake et al 1970 - don't have
%   Vogel 1970 - don't have
%   Parkhurst and Loucks 1972
%   Givnish 1979 - don't have
%   Grace et al 1980 - don't have
%   Okajima et al 2012
% leaf size (empirical): 
%   McDonald et al 2003
%   Leigh et al 2017
%   Yates et al 2010 - small leaves transpire more (but prob due also to g_s)
%   Wright et al 2017 - 
%
% effects of evaporation
%   Gates_1968 - review theory and empirical
%   Yates et al 2010 focused on evap.

% effect on photosynthesis. somewhat recent review:
%   Bernacchi et al 2009 in Photosynthesis in silico
%
% Michelatz et al 2016 - linkages to global trait spectra or something...
% Blonder and Michelatz 2018 - T_leaf vs T_air relationship (thermoregulation)

% Leigh et al 2012 NP - leaf thickness and tau
% PLoS ONE Paper I had up from guy at harbard, now davis (Z-lab)
% 
% Leigh et al 2017:
% find that effective leaf width explains cooling time constant best
% "wind speeds within rainforest canopies can regularly be < 0.5 m/s (Martin et al 1999, Stokes et al. 2006, Meinzer et al. 1995)

% term 'characteristic dimension' is from Taylor 1975

@@

% Note: this re-does Foster and Smith 1987 model but makes it easy to look at lots of gradients they did not (can I recreate some classic results?)

@@

\begin{equation}
  R_\mathrm{abs} = S_\mathrm{r} + H + L
\end{equation}

@@

% INTERESTING - TECHNICALLY, $(T_\mathrm{leaf} - T_\mathrm{air})$ factors out here...

@@

R is a fully open source programming language for statistical computing that allows users to develop their own packages with new functions. \tealeaves~takes three sets of parameter inputs: leaf parameters, environmental parameters, and physical constants (see Table \ref{table:table_input}). The package provides reasonable defaults, but users can input new values to address their question, as I demonstrate in the next section. With one or more parameter sets, \tealeaves~uses the \code{uniroot} function in R base package \pkg{stats} to find the \tleaf~that balances the leaf energy budget. It outputs the equilibrium \tleaf~and energy fluxes in a table for analysis and visualization.

@@

\begin{Schunk}
\begin{Sinput}
>   library(tealeaves)
>   # Default parameter inputs  
>   leaf_par   <- make_leafpar()
>   enviro_par <- make_enviropar()
>   constants  <- make_constants()
>   # Solve for T_leaf
>   T_leaf <- tleaf(leaf_par, enviro_par, constants, 
+                   quiet = TRUE, unitless = FALSE)
> 
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
>   library(tealeaves)
>   # Custom parameter inputs
>   leaf_par   <- make_leafpar(
+     replace = list(
+       leafsize = set_units(c(0.0025, 0.025, 0.25), "m")
+       )
+   )
>   enviro_par <- make_enviropar(
+     replace = list(
+       T_air = set_units(seq(275, 310, 5), "K")
+       )
+   )
>   constants  <- make_constants()
>   # Solve for T_leaf over a range of T_air
>   T_leaves <- tleaves(leaf_par, enviro_par, constants, 
+                       quiet = TRUE, unitless = FALSE)
> 
\end{Sinput}
\end{Schunk}
